### 2022-03-16 outputs/paper/paper.Rmd
add 2021 us gss data, data prep and work on writeup

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Roy Chan
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Introduction

# Data
The survey data that is presented within this paper is obtained from the 2021 United States General Social Survey (GSS).

## Survey methodology
The US GSS was created to collect public opinion and sentiment within the United States. The main component of the GSS is the replicating core, a central set of questions repeated in each year for the purpose of exploring trends in opinion over time. Additional modules present but not always consistently asked from year to year include topical questions relevant to the year. A major strength of the US GSS is in its long existence and replicating core, allowing tracking of contemporary sentiments over a significant period of time. Prior to 2021, surveys were administered in person resulting

Historically, the GSS has been administered in person but owing to the COVID-19 pandemic, the 2021 edition

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Supplementary survey

\newpage

# References